\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
setwd("/home/guest/Chen-Thompson-Project")

# Load your packages
library(dplyr)
library(ggplot2)
library(tidyverse)

# Set your ggplot theme
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

# Load your datasets
CO2_cleaned_rest <- read.csv("~/Chen-Thompson-Project/Data/Raw/CO2_cleaned_restructured.csv", stringsAsFactors = TRUE)
CO2_cleaned <- read.csv("~/Chen-Thompson-Project/Data/Raw/CO2_cleaned.csv", stringsAsFactors = TRUE)
QoL_Classes <- read.csv("~/Chen-Thompson-Project/Data/Raw/QoL_Classes.csv", stringsAsFactors = TRUE)
QoL_Cleaned <- read.csv("~/Chen-Thompson-Project/Data/Raw/QoL_cleaned.csv", stringsAsFactors = TRUE)

```


# Rationale and Research Questions
    #### Contains clear context for research topic
      Contains rationale for dataset of choice
      Contains one or more questions of an appropriate scope
      for the projec
Since the industrial revolution and mass production, CO2 emissions have changed a lot during the past two centuries. With the continual increasing CO2 emissions, global warming and other effects make many countries and people at a risk of lowering quality of life. 
To study the relationship between CO2 emissions and quality of life, we choose a dataset that provides the C02 levels and life qualities of countries around the world per year, which is perfectly fit for our research topic.
To begin with, we propose a question that does population size, density and development status impact CO2 emissions in worldwide(about 120 countries). Then, we try to see what is the CO2 emission trend in the US since 1750. Finally, we want to find out whether do CO2 emissions affect quality of life in worldwide.



\newpage

# Dataset Information
    ####Describes source and content of data
    Details the wrangling process from raw to processed data
    Contains a table summarizing the dataset structure
    
The dataset we use comes from The World Banks Atlas Method which uses individual countries data reported to the United Nations Framework Convention on Climate Change (UNFCCC) and other international sources. It constains information as followed: Country (Title), Code(Country), Year, CO2 emissions, Population, Area, Density, Quality of life and so on.
To get the processed data that we want, first, we input two datasets that contain information about CO2 emissions and quality of life respectively. Then, we only select the columns that we needed in year 2019 and combine these two datasets to make a clean one. For the time series analysis, we only select the CO2 emissions data in the US from 1750 to 2020 and make a new "date" column.

\newpage

# Exploratory Analysis 
    ####Flow between text and visualizations is cohesive
    Relevant exploratory information is visualized
To begin with, let's take a initial look at the dataset. In the first graph, we can see that the US had the largest CO2 emissions, with China being the second largest and Germany being the third largest. Then, we can find that there is no clear trends in relationships between CO2 emissions and population. Similarly, we cannot see any trend between CO2 emissions and population density. Finally, when we take a look at the graph about CO2 emissions and quality of life, there is again no trends.

\newpage

# Analysis
    ####Flow between text and visualizations is cohesive
    Visualizations and statistical tests pertain directly to
    specific questions
    Statistical results are reported in plain language with
    relevant statistical output in parentheses
    Findings are reported clearly in relation to research
    questions


## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




```{r echo=FALSE}
#TSA of CO2 Data in the US
CO2_US <- 
  CO2_cleaned %>%
  filter(Code == "US")%>%
  select(Country, Year, CO2.emission..Tons.)%>%
  mutate(Date = paste0(1,"-",1,"-",Year))
CO2_US$Date <- as.Date(CO2_US$Date, format = "%m-%d-%Y")

CO2_US_ts <- ts(CO2_US$CO2.emission..Tons., start = c(1750), frequency=1)
CO2_data_trend <- MannKendall(CO2_US_ts)
CO2_data_trend
summary(CO2_data_trend)
mk.test(CO2_US$CO2.emission..Tons., continuity = TRUE)

CO2_data_plot <-
ggplot(CO2_US, aes(x = Date, y = CO2.emission..Tons.)) +
  geom_point() +
  geom_line() +
  labs(title = paste("CO2 emissions in the US"),
       y="CO2.emission(Tons)",
       x="Date")+
  geom_smooth(method = lm, se = FALSE)
print(CO2_data_plot)
```

After we run the MannKendall test, we find out the z score is 23.789 and the p-value is 2.2e-16, which is much smaller than the alpha level. As a result, we could reject the null hypothesis that from 1750 to 2020, there is no CO2 emission difference in the US. And we can conclude that during the past two centuries, there is an upward trend in CO2 emissions in the US.
\newpage

# Summary and Conclusions
   ####Major findings are summarized
   Conclusions relate back to the original research context

\newpage

# References
<add references here if relevant, otherwise delete this section> 
