---
title: "1_DataVisualization"
author: "Michelle Thompson"
date: "2022-11-18"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

CO2_cleaned_rest <- read.csv("~/Chen-Thompson-Project/Data/Raw/CO2_cleaned_restructured.csv", stringsAsFactors = TRUE)
CO2_cleaned <- read.csv("~/Chen-Thompson-Project/Data/Raw/CO2_cleaned.csv", stringsAsFactors = TRUE)
QoL_Classes <- read.csv("~/Chen-Thompson-Project/Data/Raw/QoL_Classes.csv", stringsAsFactors = TRUE)
QoL_Cleaned <- read.csv("~/Chen-Thompson-Project/Data/Raw/QoL_cleaned.csv", stringsAsFactors = TRUE)

```


```{r}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

### Import your datasets

```{r Importing data}

# Relative file path (friendly for users regardless of machine)
CO2_cleaned_restructured <- read.csv("~/Chen-Thompson-Project/Data/Raw/CO2_cleaned_restructured.csv", stringsAsFactors = TRUE)
CO2_cleaned <- read.csv("~/Chen-Thompson-Project/Data/Raw/CO2_cleaned.csv", stringsAsFactors = TRUE)
QoL_Classes <- read.csv("~/Chen-Thompson-Project/Data/Raw/QoL_Classes.csv", stringsAsFactors = TRUE)
QoL_Cleaned <- read.csv("~/Chen-Thompson-Project/Data/Raw/QoL_cleaned.csv", stringsAsFactors = TRUE)

```


## EXPLORE YOUR DATASET

```{r Exploring data}
View(CO2_cleaned)
View(CO2_cleaned_restructured)
View(QoL_Classes)
View(QoL_Cleaned)


class(CO2_cleaned)
class(CO2_cleaned_restructured)
class(QoL_Cleaned)
class(QoL_Classes)


colnames(CO2_cleaned)
colnames(CO2_cleaned_restructured)
colnames(QoL_Cleaned)
colnames(QoL_Classes)

str(CO2_cleaned)
str(CO2_cleaned_restructured)
str(QoL_Cleaned)
str(QoL_Classes)

dim(CO2_cleaned)
dim(CO2_cleaned_restructured)
dim(QoL_Classes)
dim(QoL_Cleaned)

length(CO2_cleaned)
length(CO2_cleaned_restructured)
length(QoL_Classes)
length(QoL_Cleaned)
  #if we want to merge we should use co2 restructered and qol cleaned versions

head(CO2_cleaned)
head(CO2_cleaned_restructured)
head(QoL_Classes)
head(QoL_Cleaned)


summary(CO2_cleaned)  
summary(CO2_cleaned_restructured)
summary(QoL_Classes)
summary(QoL_Cleaned)

```

```{r finding NAs}

which(is.na(CO2_cleaned), arr.ind=TRUE)
which(is.na(CO2_cleaned_restructured), arr.ind=TRUE)
which(is.na(QoL_Classes), arr.ind=TRUE)
which(is.na(QoL_Cleaned), arr.ind=TRUE)
    #No NAs

```

STOP:

  To do: merge data, trim it down, make intial graphs and stats

```{r}
#Wrangling data
#Select relevant columns from co2 and Qol datasets
library(tidyverse)
CO2_cleaned_restructured <- CO2_cleaned_restructured %>% 
        rename("Percent_of_world" = "X..of.World",
               "CO2_tons" = "CO2.emission..Tons.",
               "Population.2022"= "Population.2022.",
               "Density"= "Density.km2.")

CO2_select <- select(CO2_cleaned_restructured, X, Country, Year, Population.2022, Area, Percent_of_world, Density, CO2_tons, Value)

QoL_select <- select(QoL_Cleaned, X, Country, Stability.15.., Rights.20.., Health.15.., Safety.10.., Climate.15.., Costs.15.., Popularity.10.., TotalQuality.of.life.100..)
  ##what year is QOL dataset in

##did not include ranks but I can add them subsequently

#select co2 by year
CO2_select_2019 <- 
  CO2_select %>%
  filter(Year %in% c("2019"))
  
#save as processed files
write.csv(CO2_select, row.names = FALSE, file = "~/Chen-Thompson-Project/Data/Processed/CO2_select.csv")

write.csv(QoL_select, row.names = FALSE, file = "~/Chen-Thompson-Project/Data/Processed/QoL_select.csv")

write.csv(CO2_select_2019, row.names = FALSE, file = "~/Chen-Thompson-Project/Data/Processed/CO2_select_2019.csv")

```


```{r}
#Combine CO2 2019 and QoL datasets
CO2_QoL<- merge(CO2_select_2019, QoL_select, by = "Country")
write.csv(CO2_QoL, row.names = FALSE, file = "~/Chen-Thompson-Project/Data/Processed/CO2_QoL.csv")

colnames(CO2_QoL)
#Plot initial findings
  #Co2 by Country
CO2_Country <- ggplot(CO2_QoL, aes(Country, CO2_tons)) +
    geom_bar(stat="identity", na.rm = TRUE) +
    ggtitle("CO2 Emissions (tons) by Country (2019)") +
    xlab("Country") + ylab("CO2 Emissions (tons")
 
 CO2_Country
 
  #CO2 and QoL 
CO2_QoL_relationship <- ggplot(CO2_QoL, aes(CO2_tons, TotalQuality.of.life.100..,)) +
    geom_point(stat="identity", na.rm = TRUE) +
    ggtitle("Quality of Life and CO2 Emissions") +
    xlab("CO2 Emissions") + ylab("Quality of Life")

 CO2_QoL_relationship
 
 CO2_pop_relationship <- ggplot(CO2_QoL, aes(Population.2022, CO2_tons,)) +
    geom_point(stat="identity", na.rm = TRUE) +
    ggtitle("Population and CO2 Emissions") +
    xlab("Population") + ylab("CO2 Emissions (tons)")
 
  CO2_pop_relationship
CO2_QoL$Density<- as.character(CO2_QoL$Density)
CO2_QoL$Density = substr(CO2_QoL$Density,1,nchar(CO2_QoL$Density)-4)
as.numeric(gsub(",", "", CO2_QoL$Density, fixed = TRUE))
write.csv(CO2_QoL, row.names = FALSE, file = "~/Chen-Thompson-Project/Data/Processed/CO2_QoL.csv")
  CO2_popden_relationship <- ggplot(CO2_QoL, aes(Density, CO2_tons)) +
    geom_point(stat="identity", na.rm = TRUE) +
    ggtitle("Population Density and CO2 Emissions") +
    xlab("Population Density (/km2)") + ylab("CO2 Emissions (tons)")
 
  CO2_popden_relationship
```

#developed vs developing 
```{r, fig.height = 3, fig.width = 4}
#
CO2_QOL_DEV <- read.csv("~/Chen-Thompson-Project/Data/Processed/CO2_QoL_dev.csv", stringsAsFactors = TRUE)

CO2_QOL_DEV$Density = substr(CO2_QOL_DEV$Density,1,nchar(CO2_QoL$Density))
CO2_QOL_DEV$Density <- as.numeric(gsub(",","",CO2_QOL_DEV$Density))
CO2_QOL_DEV$Density<- as.numeric(CO2_QOL_DEV$Density)

```


## Generate summary tables
12. Use the split-apply-combine strategy to generate a summary data frame. Data should be grouped by site, month, and year. Generate the mean AQI values for ozone and PM2.5 for each group. Then, add a pipe to remove instances where a month and year are not available (use the function `drop_na` in your pipe). 

13. Call up the dimensions of the summary dataset.

```{r}
#summary tables

Summary<- CO2_QOL_DEV %>%
  group_by(CO2_QOL_DEV$DevStatus)%>%
  summarise(meanCO2= mean(CO2_tons))

Summary
#graph

CO2_QOL_DEV_box <- ggplot(CO2_QOL_DEV, aes(DevStatus, CO2_tons,)) +
    geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)+
    ggtitle("Development Status and CO2 Emissions") +
    xlab("Development Status") + ylab("CO2 Emissions")
CO2_QOL_DEV_box

#remove outliers
outliers <- boxplot(CO2_QOL_DEV$DevStatus, plot=FALSE)$out


CO2_QOL_DEV_noO <- ggplot(CO2_QOL_DEV, aes(DevStatus, CO2_tons,)) +
    geom_boxplot(outlier.shape=NA)+ 
  scale_y_continuous(limits = c(0.1, 1000000000))+
    ggtitle("Development Status and CO2 Emissions") +
    xlab("Development Status") + ylab("CO2 Emissions")
  
    CO2_QOL_DEV_noO
```
```{r}
#glms

CO2.pop.glm<- lm(data = CO2_QOL_DEV, CO2_tons ~ Population.2022 + Density)
summary(CO2.pop.glm)

cor.test(CO2_QOL_DEV$CO2_tons, CO2_QOL_DEV$Population.2022)
cor.test(CO2_QOL_DEV$CO2_tons, CO2_QOL_DEV$Density)
  #population p<0.05, density not

CO2.dev.glm <- lm(data = CO2_QOL_DEV,CO2_tons ~DevStatus)
summary(CO2.dev.glm)

CO2.QOL.glm <- lm(data = CO2_QOL_DEV, TotalQuality.of.life.100.. ~ CO2_tons)
summary(CO2.QOL.glm)
```



